---
title: "Investment Analysis & Investment Tracker - A Statistical Prespective"
output: 
  pdf_document: 
    toc: true
    highlight: tango
editor_options: 
  markdown: 
    wrap: sentence
---

\newpage

**“Money that doesn’t grow at least at the pace of inflation is silently losing its worth. Beating inflation is not luck—it is a skill. Sustaining it through time is the art of investing, and compounding is the power that turns patience into wealth.”**

# **Inflation**

**what is inflation?**

Inflation is a macroeconomic phenomenon that represents the persistent and general increase in the price level of goods and services in an economy over time.
It reduces the purchasing power of money, meaning that consumers need to spend more money to purchase the same quantity of goods and services compared to an earlier period.

**Inflation is commonly measured by using the Consumer Price Index (CPI).**

## **CPI**

The Consumer Price Index (CPI) is a measure that examines the weighted average of prices of a basket of consumer goods and services, such as transportation, food, and medical care.
It is calculated by taking price changes for each item in the predetermined basket of goods and averaging them.

The CPI is used to measure inflation by tracking changes in the price level of a market basket of consumer goods and services over time.
When the CPI rises, it indicates that the average price level of goods and services has increased, which is a sign of inflation.
Conversely, when the CPI falls, it indicates deflation.

CPI values are often reported on a monthly basis, and the percentage change in CPI from one period to another is used to calculate the inflation rate.

## **CPI Over Time**

```{r echo=FALSE , fig.width = 12, fig.height = 6}
cpi_plot
```

\newpage
## **Inflation Rate Calculation**

The inflation rate is calculated using the formula:
$$
\text{Inflation Rate (\%)}=\frac{\text{CPI}_t-\text{CPI}_{t-1}}{\text{CPI}_{t-1}}\times100
$$
Where:

-    $\textbf{CPI}_\textbf{t}$ is the Consumer Price Index at the current time period.

-    $\textbf{CPI}_\textbf{t-1}$ is the Consumer Price Index at the previous time period.

**Month on Month (MoM) Inflation Rate Calculation:**
$$
\text{MoM Inflation (\%)} =
\frac{\text{CPI}_{t} - \text{CPI}_{t-1}}{\text{CPI}_{t-1}} \times 100
$$

**Year on Year (YoY) Inflation Rate Calculation:**
$$
\text{YoY Inflation (\%)} =
\frac{\text{CPI}_{t} - \text{CPI}_{t-12}}{\text{CPI}_{t-12}} \times 100
$$


## **Monthly and yearly Inflation Rate**
```{r echo=FALSE , fig.width = 14, fig.height = 18}
inflation_plot
```
